---
title: "Example 2: Semi-Confirmatory Factor Analysis"
author: "Po-Hsien Huang"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
In this example, we will show how to use `lslx` to conduct semi-confirmatory factor analysis.
The example uses data `HolzingerSwineford1939` in the package `lavaan`.
Hence, `lavaan` must be installed.

## Model Sepcification
In the following specification, `x1` - `x9` is assumed to be measurements of 3 latent factors: `visual`, `textual`, and `speed`.
```{r comment = "", message = FALSE}
equation2 <-
'
visual  :=> x1 + x2 + x3
textual :=> x4 + x5 + x6
speed   :=> x7 + x8 + x9
visual  :~> x4 + x5 + x6 + x7 + x8 + x9 
textual :~> x1 + x2 + x3 + x7 + x8 + x9 
speed   :~> x1 + x2 + x3 + x4 + x5 + x6 
visual  <=> fix(1) * visual
textual <=> fix(1) * textual
speed   <=> fix(1) * speed
'
```
The operator `:=>` means that the LHS latent factors is defined by the RHS observed variables.
In particular, the loadings are freely estimated.
The operator `:~>` also means that the LHS latent factors is defined by the RHS observed variables, but the loadings are penalized.
In this model, `visual` is mainly measured by `x1` - `x3`, `textual` is mainly measured by `x4` - `x6`, and `speed` is mainly measured by `x7` - `x9`.
However, the inclusion of the penalized loadings indicates that each measurement may not be influenced by only one latent factor. 
The operator `<=>` means that the LHS and RHS variables/factors are covaried.
If the LHS and RHS variable/factor are the same, `<=>` specifies the variance of that variable/factor.
For scale setting, `visual <=> fix(1) * visual` makes the variance of `visual` to be zero.
Details of equation syntax can be found in the section of Equation Syntax via `?lslx`.

## Object Initialization
`lslx` is written as an `R6` class.
Everytime we conduct analysis with `lslx`, an `lslx` object must be initialized.
The following code initializes an `lslx` object named `r6_lslx2`.
```{r comment = "", message = FALSE}
library(lslx)
r6_lslx2 <- lslx$new(equation = equation2,
                     sample_data = lavaan::HolzingerSwineford1939)
```
Here, `lslx` is the object generator for `lslx` object and `new` is the build-in method of `lslx` to generate a new `lslx` object.
The initialization of `lslx` requires users to specify a equation for model specification (argument `equation`) and a data set to be fitted (argument `sample_data`).
The data set must contains all the observed variables specified in the given equation.
In is also possible to initialize an `lslx` object via sample moments (see Example 3).

It is possible to see the model specification via the `extract_specification` method.
```{r comment = "", message = FALSE}
r6_lslx2$extract_specification()
```
The row names show the coefficient names. 
The most two relevant columns are `type` which shows the type of the coefficient and `start` which shows the starting value.
In `lslx`, many `extract` methods are defined that can be used to extract information stored in `lslx` object.
For details, please see the the section of Extract-Related Methods via `?lslx`.

## Model Fitting
After an `lslx` object is initialized, method `fit` can be used to fit the specified model into the given data.
```{r comment = "", message = FALSE}
r6_lslx2$fit(penalty_method = "mcp",
             lambda_grid = seq(.05, .30, .05),
             gamma_grid = c(5, 10))
```
The fitting process requires users to specify the penalty method (argument `penalty_method`) and the considerd penalty levels (argument `lambda_grid` and `gamma_grid`).
In this example, the `mcp` penalty is implemented on the lambda grid `c(.05, .10, .15, .20, .25, .30)` and gamma grid `c(5, 10)`.
Note that in this example `lambda = 0` is not considered because it may result in unidentified result.
All the fitting result will be stored in the `fitting` field of `r6_lslx2`.


## Model Summarizing
Unlike traditional SEM analysis, `lslx` fit the model into data under all the penalty levels considered.
To summarize the fitting result, a selector to determine an optimal penalty level must be specified.
Availble selectors can be found in the section of Penalty Level Selection via `?lslx`.
The following code summarize the fitting result under the penalty level selected by Bayesian information criterion.
```{r comment = "", message = FALSE, fig.width = 24, fig.height = 14}
r6_lslx2$summarize(selector = "bic")
```
In ths example, we can see that most penalized parameters are zero under the selected penalty level except for `x9<-visual`, which shows the benefit of using the semi-confirmatory approach.
So far, the hypothesis testing for the model coefficient are based on the expected Fisher information and doesn't consider the presence of model selection.
In the future, other standard error formula and valid post model selection inference will be available.


## Visualization
`lslx` provides three methods for visualizaing the fitting result.
The method `plot_numerical_condition` shows the numerical condition under all the penalty levels.
The following code plots the values of `n_iter_out` (number of iterations in outer loop), `objective_gradient_abs_max` (maximum of absolute value of gradient of objective function), and `objective_hessian_convexity` (minimum of univariate approximate hessian).
The plot can be used to evaluate the quality of numerical optimization.
```{r comment = "", message = FALSE, fig.width = 8, fig.height = 4, dpi=300, out.width=600, out.height=300}
r6_lslx2$plot_numerical_condition(criterion = c("n_iter_out", 
                                                "objective_gradient_abs_max",
                                                "objective_hessian_convexity"))
```


The method `plot_goodness_of_fit` shows the values of information criteria and fit indices under all the penalty levels.
The following code plots the values of `loss` (ML discrepancy function value evaluated at the current estimate), `aic`, and `bic`.
```{r comment = "", message = FALSE, fig.width = 8, fig.height = 4, dpi=300, out.width=600, out.height=300}
r6_lslx2$plot_goodness_of_fit(selector = c("loss", "aic", "bic"))
```


The method `plot_coefficient` shows the solution path of coefficients in the given block.
The following code plots the solution paths of all coefficients in the block `y<-f`, which contains all the regression coeffcients from latent factors to observed variables (i.e., factor loadings).
```{r comment = "", message = FALSE, fig.width = 8, fig.height = 4, dpi=300, out.width=600, out.height=300}
r6_lslx2$plot_coefficient(block = "y<-f")
```
