---
title: "Semi-Confirmatory Factor Analysis with Missing Data"
author: "Po-Hsien Huang"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r comment = "", message = FALSE, setup, include=FALSE}
options(digits = 3)
options(width = 100)
```

In this example, we will show how to use `lslx` to conduct semi-confirmatory factor analysis.
The example uses data `HolzingerSwineford1939` in the package `lavaan`.
Hence, `lavaan` must be installed.

## Missing Data Construction
Because `HolzingerSwineford1939` doesn't contain missing values, we use the code in `semTools` to create `NA` (see the example of `twostage` function in `semTools`).
```{r comment = "", message = FALSE}
data <- lavaan::HolzingerSwineford1939
data$x5 <- ifelse(data$x1 <= quantile(data$x1, .3), NA, data$x5)
data$age <- data$ageyr + data$agemo/12
data$x9 <- ifelse(data$age <= quantile(data$age, .3), NA, data$x9)
```
By the construction, we can see that the missingness of `x5` depends on the value of `x1` and the missingness of `x9` relies on the `age` variable. Note that `age` is created by `ageyr` and `agemo`. 
Since `ageyr` and `agemo` are not the variables that we are interested, the two variables are treated as auxiliary in the later analysis.

## Model Sepcification and Object Initialization
The following model specification is the same to our example of semi-confirmatory factor analysis (see `vignette("factor-analysis")`).
```{r comment = "", message = FALSE}
model <-
'
visual  :=> x1 + x2 + x3
textual :=> x4 + x5 + x6
speed   :=> x7 + x8 + x9
visual  :~> x4 + x5 + x6 + x7 + x8 + x9 
textual :~> x1 + x2 + x3 + x7 + x8 + x9 
speed   :~> x1 + x2 + x3 + x4 + x5 + x6 
visual  <=> fix(1) * visual
textual <=> fix(1) * textual
speed   <=> fix(1) * speed
'
```
To initialize an `lslx` object with auxiliary variables, we need to specify the `auxiliary_variable` argument.
`auxiliary_variable` argument only accepts numeric variables.
If any categorical variable ia considered as a valid auxiliary variable, user should transform it as a set of dummy variables first. One possible method is using `model.matrix` function.
```{r comment = "", message = FALSE}
library(lslx)
r6_lslx <- lslx$new(model = model,
                    data = data,
                    auxiliary_variable = c("ageyr", "agemo"))
```

## Model Fitting
So far, the specified auxiliary variables are only stored in `lslx` object. 
They are actually used after implementing the `fit` related methods.
```{r comment = "", message = FALSE}
r6_lslx$fit(penalty_method = "mcp",
            lambda_grid = seq(.02, .30, .02),
            delta_grid = c(5, 10))
```
By default, `fit` related methods implement two-step method (possibly with auxiliary variables) for handling missing values.
User can specify the missing method explicitly via `missing_method` argument.
Another missing method in the current version is listwise deletion.
However, listwise deletion has no theoretical advantages over the two step method.

## Model Summarizing
The following code summarizes the fitting result under the penalty level selected by Bayesian information criterion (BIC).
The `number of missing pattern` shows how many missing patterns present in the data set (include the complete pattern).
If the `lslx` object is initialized via raw data, by default, a corrected sandwich standard error will be used for coefficient test.
The correction is based on the asymptotic covariance of saturated derived by full information maximum likelihood.
Also, the mean adjusted likelihood ratio test is based on this quantity.
For the reference, please see the section of Missing Data in `?lslx`
```{r comment = "", message = FALSE, fig.width = 24, fig.height = 14}
r6_lslx$summarize(selector = "bic")
```
